A Simple Minibatch Acceptance Test for MCMC

### 2016-05-16 nips_2016.tex
add figs
updates to abstract

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

@@

\title{A New Metropolis-Hastings Test for Fast Minibatch MCMC Inference and Optimization}

@@

\begin{abstract}
Markov chain Monte Carlo (MCMC) methods are one of the most popular techniques in machine learning
for Bayesian inference. With large datasets, however, MCMC methods can be intractably slow due to
the need to compute the likelihood of all data points each iteration. Borrowing from the
optimization field, one way to make MCMC faster is to compute the posterior based on a random subset
of the data each iteration. Such stochastic MCMC has received increased research attention in recent
years, but these methods generally require adaptively increasing the minibatch size each iteration,
which can nullify the benefits if too much data is needed. In addition, with large datasets and
model optimization, it is unclear if obtaining a full posterior distribution is desirable over
getting a high-quality point MAP or ML estimate of the parameters.  In this paper, we present a new
Metropolis-Hastings test to fit in our alternative perspective on stochastic MCMC methods for big
data model optimization. We demonstrate the benefits of our approach with theoretical and empirical
results.
\end{abstract}

@@

Markov chain Monte Carlo (MCMC) sampling is one of the most popular techniques for Bayesian
computation. MCMC algorithms propose samples from a proposal distribution $q$, and decide whether to
accept or reject them based on a Metropolis-Hastings test~\cite{Metropolis1953,hastings70}. With
traditional Bayesian posterior inference, the posterior distribution $p(\theta \mid x_1, \ldots,
x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ of parameter $\theta$ based on conditionally
independent data $\{x_i\}_{i=1}^n$ is the target distribution which one aims to approximate by
sampling $\theta$ values. When $n$ is large (e.g., in the billions), it is computationally difficult
to compute the exact posterior because the likelihood must be evaluated on each $x_i$.

@@

In this experiment, we also illustrate that by appling the temperature, we can achieve slight faster convergence than the constant temperature one.
\begin{figure}[ht]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  %\includegraphics[width=0.5\linewidth]{empty}
  \caption{This will be a sequence of images for logistic regression.}
\end{figure}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

@@

\title{A Simple Minibatch Acceptance Test for MCMC}

@@

\begin{abstract}
Markov chain Monte Carlo (MCMC) methods have many applications in
machine learning. We are particularly interested in their application
to modeling very large datasets, where it is impractical to
perform Metropolis-Hastings tests on an entire dataset. Instead we
describe a novel acceptance test which is applied to minibatches of
data. Previous work on minibatch Metropolis-Hastings use adaptive
batch sizes, which may use far more than one minibatch of data and/or
yield few samples.  Here we describe a simpler test on fixed-sized
minibatches. Instead of adaptive batch size, this test requires a
precondition on the variance of the log proposal probability
ratio. Given a measurement of this variance, minibatch size can be
chosen to support the test in one step, and yielding a number of
samples proportional to the batch/minibatch ratio. Alternatively the
variance condition can be used to adjust the temperature of the
sampling distribution to a ``natural'' value to provide regular
samples at a given minibatch size. The resulting proposal/test is only
slightly more complex than a simple SGD update. In this paper we
derive the test, discuss its implementation, and present several
experiments.

\end{abstract}

@@

Markov chain Monte Carlo (MCMC) sampling is a powerful method for
computation on intractable distributions. MCMC algorithms propose
samples from a proposal distribution $q$ which is in general different
from the target distribution $p$, and decide whether to accept or
reject them based on an acceptance test. The acceptance test is
usually a Metropolis-Hastings
test~\cite{Metropolis1953,hastings70}. For example, for Bayesian
posterior inference, the posterior distribution $p(\theta \mid x_1,
\ldots, x_N) \propto p(\theta)\prod_{i=1}^np(x_i \mid \theta)$ of
parameter $\theta$ based on conditionally independent data
$\{x_i\}_{i=1}^n$ is the target distribution which one aims to
approximate by sampling $\theta$ values. When $n$ is large (e.g., in
the billions), it is computationally difficult to compute the exact
posterior because the likelihood must be evaluated on each $x_i$.

@@

In this experiment, we also illustrate that by appling the temperature, we can achieve slight faster convergence than the constant temperature one.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/acc_hyc.png}
        \caption{prediction accuracy}
        \label{fig:logistic_pa}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/ll_hyc.png}
        \caption{loglikelihood}
        \label{fig:logistic_ll}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/roofline_hyc.png}
        \caption{roofline property}
        \label{fig:logistic_roof}
    \end{subfigure}
    \caption{Experiments of logistic regression}\label{fig:logistic}
\end{figure}

\begin{figure}[ht]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  %\includegraphics[width=0.5\linewidth]{empty}
  \caption{This will be a sequence of images for logistic regression.}
\end{figure}
